---
title: "Compute DTP score for a sample"
author: "Chaitra Sarathy"
date: "2023-12-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries 

```{r}
library(tidyverse)
library(survival)
library(survminer)
library(ggrepel)
library(sva)
```

Load annotation gtf file to map gene ids to symbols

```{r}
HUMAN_GTF_FILE_PATH = "gencode.v31.annotation.gtf"
features <- rtracklayer::import(HUMAN_GTF_FILE_PATH)
features_gene = as.data.frame(features)
features_gene = features_gene[which(features_gene$type == "gene"),]
keep <- c("gene_id", "gene_name")
features_gene <- features_gene[, names(features_gene) %in% keep]
rownames(features_gene) <- features_gene$gene_id
```

Define file names (as processed from OICR)

```{r}
# sample_counts_star <- "DTP_0002_Co_P_WT_1I.genes.results"
# sName <- "OICR-CRC-002" 

sample_counts_star <- "DTP_0003_Li_P_WT_2P.genes.results"
sName <- "OICR-CRC-004" 
```

Import sample counts (as processed from OICR)

```{r}
s1 <- read.table(file = sample_counts_star, header = T)
```

Merge ids and symbols, collapse duplicated symbols

```{r}
# summarise expression for duplicated genes

t1 <- s1 %>% dplyr::left_join(features_gene, "gene_id")

exp.collapsed.mean <- t1 %>% dplyr::select(expected_count, TPM, FPKM, gene_name) %>%
  na.omit() %>%
  group_by(gene_name) %>%
  summarise_all(mean) %>%
  tibble::column_to_rownames('gene_name') %>%
  as.matrix()

head(rownames(exp.collapsed.mean))

dim(exp.collapsed.mean)

nrow(exp.collapsed.mean)

sTab <- exp.collapsed.mean %>% as.data.frame() %>% dplyr::select(FPKM)
colnames(sTab) <- sName
```

Read autophagy genes and load scoring functions
```{r}
gmtFile <- "geneset_autophagy.gmt"
gs.autophagy <- clusterProfiler::read.gmt(file.path(getwd(), "geneset", gmtFile))
autoGenes <- toupper(gs.autophagy$gene)

source("0b_gsscoring_functions.R")
source("0c_survival_functions.R")
```

Load TCGA matrix & merge new samples with count matrix from TCGA
```{r}
load("TCGA_COADREAD_GDC_annscaledcounts.RData")
dim(tcga_coadread_ann)

cgenes <- intersect(rownames(tcga_coadread_ann), rownames(sTab))

mer.mat <- cbind(tcga_coadread_ann[rownames(tcga_coadread_ann) %in% cgenes,],
                 sTab[rownames(sTab) %in% cgenes,])

colnames(mer.mat)[ncol(mer.mat)] <- sName
```

batch correction

```{r}
library(sva)
batch = c(rep("tcga", ncol(tcga_coadread_ann)),
            rep("OICR", ncol(sTab)))
# reference-batch version, with covariates
combat_edata = ComBat(dat=mer.mat, batch=batch, mod=NULL)
```

zscore
```{r}
# mer.mat.zscored <- t(scale(t(mer.mat)))
mer.mat.zscored <- t(scale(t(combat_edata)))
tail(colnames(mer.mat.zscored))
```

compute score
```{r}
cohort = "Merged: TCGA-COAD-READ with Patient sample"
autoScoreTab.go.wt <- compAutophagyScore(data.exprs.norm.z = mer.mat.zscored,
                                      autoGenes = autoGenes, 
                                      scoreType = "norm",
                                      datasetID = cohort)
autoScoreTab.go.wt$ap.score <- round(autoScoreTab.go.wt$ap.score, digits = 3)
```

```{r, fig.height=7, fig.width=7}
s1 <- survivalAnalysis(survTab = survTab,
                               sigScoreTab = autoScoreTab.go.wt,
                               type = "autophagy", 
                               method = "n",
                               plotTitle = paste("TCGA: Scaled with ", sName, " (inherent)", sep=": "))

s2 <- survivalAnalysis(survTab = survTab, 
                               sigScoreTab = autoScoreTab.go.wt, 
                               type = "autophagy", 
                               method="xx",
                               plotTitle = paste("TCGA: Scaled with ", sName, " (applied)", sep=": "),
                               cutType="manual", 
                               cutVal=0.126)
s1$surv.plot
s2$surv.plot
```

Plot score of OICR sample

```{r, fig.height=7, fig.width=10}
autoScoreTab.go.wt$ind = 1:nrow(autoScoreTab.go.wt)
autoScoreTab.go.wt %>% ggplot(aes(y=ap.score, x=ind, label=sample)) +
  geom_point(size=3, color=ifelse(autoScoreTab.go.wt$sample==sName, "red", "#006837"), alpha=0.6)+
  geom_hline(yintercept = 0.126, size=1, col="red")+
  xlab("") +
  ylab(paste("Scores (TCGA samples with ", sName, " )", sep="")) +
  geom_label_repel(aes(label = ifelse(autoScoreTab.go.wt$sample==sName, sName,'')),
                  box.padding   = 0.5, 
                  segment.color = 'grey50',
                  max.overlaps = Inf)+
  theme_classic() +
  theme(plot.title = element_text(size=24, color="black"),
        axis.title = element_text(size = 12, color="black"),
        axis.text.x = element_text(angle = 0, size=12, hjust=1, color="black"),
        axis.text.y = element_text(angle = 0, size=12, hjust=1, color="black"),
        axis.title.x = element_text( size=14,  color="black"),
        axis.title.y = element_text(size=14,  color="black"))
```



